

```{r setup, include=FALSE, echo=FALSE, message=FALSE}

# Required packages
library(Hmisc)
library(kableExtra)
library(tidyverse)
library(here)
library(skimr)
library(gmodels)
library(gtsummary)
library(rpart)
library(mice)
library(rms)
library(glmnet)
library(cowplot)
library(hrbrthemes)
library(pROC)
library(ResourceSelection)


# Ser global markdown options
options(prType='plain')
options(digits=3)
knitrSet(lang='markdown', h=6, w=6)

knitr::opts_chunk$set(
  fig.path = "images/", 
  cache.path = 'cache/images-',
  fig.align = "center",
  warning = FALSE,
  message = FALSE
)

# Set global gtsmmary options
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 'small', data_row.padding = gt::px(1))")
```

`r hidingTOC(buttonLabel="Outline", levels=2)`

```{r}
# Load dataset
load(file="/Users/Lauren/Box/RePORT Brazil/Datasets/Save/report_dissertation_lsp.Rdata") # original dataset with misssing
load(file="/Users/Lauren/Box/RePORT Brazil/Datasets/Save/imputed_analysis_data.Rdata") # mi_sum dataset based on summarized imputation
load(file="/Users/Lauren/Box/RePORT Brazil/Datasets/Save/imputed_analysis_data_long.Rdata") # mi_long dataset based on summarized imputation

# Load functions
source(here("Manuscript", "manuscript_functions.R"))

# ----- SETUP -----

# Seed for all analyses
SEED = 1408

# Save outcome vector 
outcome_vector <- mi_sum$unsuccessful
```


# RCS for Age


```{r}
# Model variables with outcome variable first
model_vars <- Cs(unsuccessful, hiv, diabetes_yn, lab_hgb, evertb, xray_cavit, smear_pos, rcs(age,4), female, bmi, dishx_any_minus, educ_years, alcoholhx, drughx, smokhx, non_white)

length_model_vars <- length(model_vars)

# Model formula 
model_formula <- as.formula(paste0("unsuccessful ~", (paste(model_vars[2:length_model_vars], collapse = "+"))))
```


## Model selection

```{r, cache=TRUE}
set.seed(SEED)
seed=SEED

# Approximate model 
approx_full_formula <- as.formula(paste0("predvals_full_model ~", (paste(model_vars[2:length_model_vars], collapse = "+"))))

# Set data distribution
dd <- suppressWarnings(datadist(mi_sum))
options(datadist="dd")

save_results_all <- model_development(data=mi_sum, 
                                  outcome=mi_sum$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=500,                                    
                                  threshold=60,
                                  seed=SEED)
```


```{r}
save_results_all$raw_overview
```


## Model comparison

### Model coefficients
```{r}
save_results_all$coefficients_table
```


### Model performance
```{r}
save_results_all$models_performance
```


## Boot model 

### C-statistics
```{r}
save_results_all$c_statistics
```


Other model performace measures

```{r}
boot_form <- as.formula(unsuccessful ~ lab_hgb + hiv + drughx + diabetes_yn + rcs(age,4) + +educ_years + smokhx + female + bmi + smear_pos + 
    non_white)

boot_model <- lrm(boot_form, data = mi_sum, x = T, y = T)

boot_val <- validate(boot_model, B=200, seed=SEED)

boot_val
```


### Calibration Plot
```{r}
save_results_all$boot_calibration
```


The calibration curve from Frank Harrell's package is: 
```{r}
# Calibration curve
plot(calibrate(save_results_all$boot_model, B=200), xlab="Predicted Probability")
```


### ROC curve
```{r}
roc <- save_results_all$roc_shrink

plot(roc)
plot(ci.thresholds(roc), type="shape", col="lightblue")
plot(roc, add=TRUE)
```



```{r}
# ---- Save to Excel ---
overview <- save_results_all$raw_overview
coefficients <- save_results_all$coefficients_table
performance <- save_results_all$models_performance

library(openxlsx)

# Create a blank workbook
results <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(results, "Model-Selection")
addWorksheet(results, "Coefficients")
addWorksheet(results, "Performance")

# Write the data to the sheets
writeData(results, sheet = "Model-Selection", x = overview)
writeData(results, sheet = "Coefficients", x = coefficients)
writeData(results, sheet = "Performance", x = performance)

# Export the file
saveWorkbook(results, here("Manuscript", "tables", "sensitivity_analyses_results.xlsx"), overwrite=TRUE)
```



# CC and Imputataion  

```{r}
# Create age groups (original data with missing)
data <- data %>% 
  mutate(age_group = fct_case_when(
    age < 25 ~ "<25",
    age >= 25 & age <35 ~ "25-35",
    age >=35 & age <45 ~ "35-45",
    age >=45 & age <55 ~ "45-55",
    age >= 55 ~ "55+"
  ))


# Create age groups (mi_long)
mi_long <- mi_long %>% 
  mutate(age_group = fct_case_when(
    age < 25 ~ "<25",
    age >= 25 & age <35 ~ "25-35",
    age >=35 & age <45 ~ "35-45",
    age >=45 & age <55 ~ "45-55",
    age >= 55 ~ "55+"
  ))
```



```{r}
# Model variables with outcome variable first (original model vars with age_group)
model_vars <- Cs(unsuccessful, hiv, diabetes_yn, lab_hgb, evertb, xray_cavit, smear_pos, age_group, female, bmi, dishx_any_minus, educ_years, alcoholhx, drughx, smokhx, non_white)

length_model_vars <- length(model_vars)

# Model formula 
model_formula <- as.formula(paste0("unsuccessful ~", (paste(model_vars[2:length_model_vars], collapse = "+"))))

# Approximated model formula (predicting predicted values)
approx_full_formula <- as.formula(paste0("predvals_full_model ~", (paste(model_vars[2:length_model_vars], collapse = "+"))))

# Seed for all analyses
SEED = 5643
```


Repeat all model development steps within complete case and each imputed dataset

## Complete case

```{r, cache=TRUE}
complete_case <- mi_long %>%  filter(.imp==1 & complete_case==1)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(complete_case))
options(datadist="dd")

save_results_complete_case <- model_development(data=complete_case, 
                                  outcome=complete_case$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)

save_results_complete_case$`raw_overview `

```

## Imputatation 1 


```{r, cache=TRUE}
imputation1 <- mi_long %>%  filter(.imp==1)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation1))
options(datadist="dd")

save_results_imputation1 <- model_development(data=imputation1, 
                                  outcome=imputation1$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)
```


## Imputatation 2


```{r, cache=TRUE}
imputation2 <- mi_long %>%  filter(.imp==2)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation2))
options(datadist="dd")

save_results_imputation2 <- model_development(data=imputation2, 
                                  outcome=imputation2$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)

```


## Imputatation 3


```{r, cache=TRUE}
imputation3 <- mi_long %>%  filter(.imp==3)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation3))
options(datadist="dd")

save_results_imputation3 <- model_development(data=imputation3, 
                                  outcome=imputation3$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)
```


## Imputatation 4


```{r, cache=TRUE}
imputation4 <- mi_long %>%  filter(.imp==4)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation4))
options(datadist="dd")

save_results_imputation4 <- model_development(data=imputation4, 
                                  outcome=imputation4$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60, 
                                  seed=SEED)

```


## Imputatation 5


```{r, cache=TRUE}
imputation5 <- mi_long %>%  filter(.imp==5)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation5))
options(datadist="dd")

save_results_imputation5 <- model_development(data=imputation5, 
                                  outcome=imputation5$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)
```


## Imputatation 6


```{r, cache=TRUE}
imputation6 <- mi_long %>%  filter(.imp==6)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation6))
options(datadist="dd")

save_results_imputation6 <- model_development(data=imputation6, 
                                  outcome=imputation6$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)

```


## Imputatation 7


```{r, cache=TRUE}
imputation7 <- mi_long %>%  filter(.imp==7)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation7))
options(datadist="dd")

save_results_imputation7 <- model_development(data=imputation7, 
                                  outcome=imputation7$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)

```


## Imputatation 8


```{r, cache=TRUE}
imputation8 <- mi_long %>%  filter(.imp==8)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation8))
options(datadist="dd")

save_results_imputation8 <- model_development(data=imputation8, 
                                  outcome=imputation8$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)

```


## Imputatation 9


```{r, cache=TRUE}
imputation9 <- mi_long %>%  filter(.imp==9)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation9))
options(datadist="dd")

save_results_imputation9 <- model_development(data=imputation9, 
                                  outcome=imputation9$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)
```


## Imputatation 10


```{r, cache=TRUE}
imputation10 <- mi_long %>% filter(.imp==10)
                                   
# Set data distribution
dd <- suppressWarnings(datadist(imputation10))
options(datadist="dd")

save_results_imputation10 <- model_development(data=imputation10, 
                                  outcome=imputation10$unsuccessful, 
                                  model_formula=model_formula, 
                                  approx_full_formula=approx_full_formula, 
                                  bootstraps=200, 
                                  threshold=60,
                                  seed=SEED)
```


## Comparison

```{r}
results <- list(
  complete_case = save_results_complete_case,
  imputation1 = save_results_imputation1,
  imputation2 = save_results_imputation2, 
  imputation3 = save_results_imputation3, 
  imputation4 = save_results_imputation4, 
  imputation5 = save_results_imputation5, 
  imputation6 = save_results_imputation6, 
  imputation7 = save_results_imputation7, 
  imputation8 = save_results_imputation8, 
  imputation9 = save_results_imputation9, 
  imputation10 = save_results_imputation10
     )

col_names <- Cs(variable, complete_case, imputation1, imputation2, imputation3, imputation4, imputation5, imputation6, imputation7, imputation8, imputation9, imputation10)

# Full model variables and coefficients 
full_models <- lapply(results, function(x) x$coefficients_table[Cs(variable, full_model)])  %>%  reduce(left_join, by = "variable")
colnames(full_models) <- col_names

# Selected model variables and coefficients 
selected_models <- lapply(results, function(x) x$coefficients_table[Cs(variable, selected_model)])  %>%  reduce(left_join, by = "variable")
colnames(selected_models) <- col_names

# Boot medians 
boot_medians <- lapply(results, function(x) x$coefficients_table[Cs(variable, coef_boot_median)])  %>%  reduce(left_join, by = "variable")
colnames(boot_medians) <- col_names

# boot model variables and coefficients  
boot_models <- lapply(results, function(x) x$coefficients_table[Cs(variable, coef_boot_model)])  %>%  reduce(left_join, by = "variable")
colnames(boot_models) <- col_names

# Shrunken coefficients (from boot model) 
shrink_models <- lapply(results, function(x) x$coefficients_table[Cs(variable, coef_shrink)]) %>%  reduce(left_join, by = "variable")
colnames(shrink_models) <- col_names
```


### AUC from full and boot models

```{r}
# AUCs
auc_full <- lapply(results, function(x) x$c_statistics[Cs(full_model, boot_model)]) 
auc_full                                 
```


### ROC curves 
```{r}
plot(save_results_complete_case$roc_shrink)
plot(save_results_imputation1$roc_shrink, add=TRUE, col=3, lty=2)
plot(save_results_imputation2$roc_shrink, add=TRUE, col=4, lty=2)
plot(save_results_imputation3$roc_shrink, add=TRUE, col=5, lty=2)
plot(save_results_imputation4$roc_shrink, add=TRUE, col=6, lty=2)
plot(save_results_imputation5$roc_shrink, add=TRUE, col=7, lty=2)
plot(save_results_imputation6$roc_shrink, add=TRUE, col=8, lty=2)
plot(save_results_imputation7$roc_shrink, add=TRUE, col=9, lty=2)
plot(save_results_imputation8$roc_shrink, add=TRUE, col=10, lty=2)
plot(save_results_imputation9$roc_shrink, add=TRUE, col=11, lty=2)
plot(save_results_imputation10$roc_shrink, add=TRUE, col=12, lty=2)
```


### Coefficients

#### Full models

```{r}
full_models
```

#### Boot 50 models

```{r}
boot_models
```

#### Selected models

```{r}
selected_models
```

#### Shrunken coefficients 

```{r}
shrink_models
```


```{r}
library(openxlsx)

# Create a blank workbook
imputation_results <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(imputation_results, "Full model coefficients")
addWorksheet(imputation_results, "Selected coefficients")
addWorksheet(imputation_results, "Bootstrap coefficients")
addWorksheet(imputation_results, "Boot model coefficients")
addWorksheet(imputation_results, "Boot-shrink coefficients")

# Write the data to the sheets
writeData(imputation_results, sheet = "Full model coefficients", x = full_models)
writeData(imputation_results, sheet = "Selected coefficients", x = selected_models)
writeData(imputation_results, sheet = "Bootstrap coefficients", x = boot_medians)
writeData(imputation_results, sheet = "Boot model coefficients", x = boot_models)
writeData(imputation_results, sheet = "Boot-shrink coefficients", x = shrink_models)

# Export the file
saveWorkbook(imputation_results, here("Manuscript", "tables", "imputation_results.xlsx"), overwrite=TRUE)
```

# External validation

External validation of Costa-Veiga model

` x = exp[log(0.72)*(age 25-44) + log(1.39)*(age 45-64) + log(4.37)*(age >= 65) + log(4.93)*(HIV) + log(2.09)*(other disease comorbidities) + log(2.29)*(IV drug use) + log(1.44)*(retreatment case)] `

```{r}
cv <- mi_sum %>% 
  mutate(
    age_group = case_when(
      age <25 & age>15 ~ "15-24",
      age <45 & age>=25 ~ "25-44",
      age <65 & age>=45 ~ "45-64",
      age >=65 ~ "65+",
      TRUE ~ NA_character_),
    diabetes = diabetes_yn,
    current_drug_use = ifelse(drughx == "Current", 1, 0),
    comorbidity = dishx_any_minus)

analysis <- cv %>% 
  mutate(lp = 
           log(0.72)*(age_group=="25-44") +
           log(1.39)*(age_group=="45-64") + 
           log(4.37)*(age_group=="65+") + 
           log(4.93)*(hiv==1) + 
           log(2.09)*(comorbidity==1) + 
           log(2.29)*(current_drug_use==1) + 
           log(1.44)*(evertb==1),
         prob_outcome = 1/(1+exp(-lp)))
  
```


## Original model

```{r}
# Use linear predictor as offset term (no slope or intercept updating)
ev_glm(method="original", lp="lp", outcome="unsuccessful", data=analysis, samples=200, return="performance")
```


## Update intercept

```{r}
# Update intercept
ev_glm(method="update intercept", lp="lp", outcome="unsuccessful", data=analysis, samples=200, return="performance")
```


## Model recalibration 

```{r}
# Update slope and intercept 

# With GLM
ev_glm(method="model recalibration", lp="lp", outcome="unsuccessful", data=analysis, samples=200, return="model")
ev_glm(method="model recalibration", lp="lp", outcome="unsuccessful", data=analysis, samples=200, return="performance")
ev_glm(method="model recalibration", lp="lp", outcome="unsuccessful", data=analysis, samples=200, return="cal_plot")


# With LRM 
ev_lrm(method="model recalibration", lp="lp", outcome="unsuccessful", data=analysis, samples=200, boots=50, return="c_stat")
```


## Model revision

```{r}
# Re-fit with predictors
formula <- as.formula(unsuccessful ~ female + age_group + hiv + comorbidity + current_drug_use + evertb)

# With GLM
ev_glm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=200, return="model")
ev_glm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=200,return="performance")
ev_glm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=1, return="cal_plot")
roc <- ev_glm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=200, return="roc")

roc 
plot(roc)

# With LRM 
ev_lrm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=200, boots=50, return="c_stat")
ev_lrm(method="model revision", lp="lp", outcome="unsuccessful", formula=formula, data=analysis, samples=200, boots=1, return="cal_plot")
```


